\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{cite}
\usepackage{amsmath,amssymb}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{url}
\usepackage{hyperref}
\hypersetup{hidelinks}

\begin{document}

\title{JMCP V1: A Local, Production-Shaped Control Plane with\\
Cryptographic Approval and Conformance-Gated\\
Multi-Agent Development}

\author{\IEEEauthorblockN{JMCP Governance Slice}
\IEEEauthorblockA{\textit{Joint Master Control Plane Project}\\
veox.ai\\
jepson@veox.ai}}

\maketitle

\begin{abstract}
We present JMCP V1, a single-machine control plane for protocol-governed
automation that preserves production-grade boundaries without depending on
remote services. JMCP is the system; JCP/1.0.0 is the message contract; and
the JMCP Plane Control Mechanism (JPCM) is the delivery and replay profile
carried by an in-process, replayable event bus. The implementation is a Rust
workspace of twelve library crates and five binaries (approximately 6{,}800
lines of Rust under 83 automated tests), a React/TypeScript operator cockpit,
and a Rust terminal interface. Every shared message is normalized to a signed
JCP/1.0.0 envelope before it enters routing, approval, audit, replay, or
execution. We make three contributions. First, an Ed25519 approval-token
scheme with a race-safe local key store that gives single-use, expiring,
forgery-resistant approvals while keeping audit records secret-safe. Second, a
bounded adapter model that delegates real work to external engines---a Jekko
worker engine reached over HTTP and an OpenAI-compatible fusion endpoint, plus
a Jeryu ecosystem producer that drives a live tool-relationship graph---without
ceding policy authority. Third, and most novel, a conformance-gated
development process in which two autonomous agents collaborate against a shared
tree under a regression ratchet: a static auditor (Jankurai) scores the
repository, and a pre-commit/pre-push hook plus a CI lane reject any change
that lowers the score or introduces a new cap or finding. We describe the
architecture, the security model, the reproducibility standard (CI--local
parity), and an evaluation of the ratchet's ability to block regressions while
permitting net improvements.
\end{abstract}

\begin{IEEEkeywords}
control plane, protocol governance, Ed25519, approval workflow, event replay,
static conformance, multi-agent systems, continuous integration,
reproducibility
\end{IEEEkeywords}

\section{Introduction}

Automation systems that act on behalf of a user---routing work, calling
external tools, mutating state, and spending money---are usually justified by
remote infrastructure: a cloud broker, a managed queue, a hosted policy engine.
JMCP V1 takes the opposite position. Its claim is that a single-machine
deployment can still preserve production-grade boundaries: durable state,
protocol validation, explicit approval, replayable events, auditable decisions,
and deterministic tests, all without requiring remote services. Local execution
is the default operating profile, not a demo mode.

This stance matters because the trustworthy parts of a control plane are
exactly the parts that are tempting to skip when ``it is only running
locally.'' If approvals are advisory, if logs leak secrets, if a replay re-runs
side effects, or if tests pass only because a developer's machine happened to
have warm state, then the local system is not a smaller production system---it
is an unaudited one. JMCP V1 is designed so that the local profile is the
audited profile.

We also treat the \emph{construction} of the system as a first-class concern.
JMCP V1 was built by two autonomous coding agents working concurrently against
one repository. That setting makes conformance regressions easy: one agent's
fix can silently re-introduce a finding another agent removed. We therefore
make non-regression a mechanical property of the repository rather than a
matter of reviewer vigilance.

\noindent\textbf{Contributions.}
\begin{enumerate}
\item A protocol-first local architecture in which every shared message is a
signed JCP/1.0.0 envelope, carried by an in-process replay bus (JPCM), with the
Rust backend as the sole policy authority (\S\ref{sec:arch}).
\item A cryptographic approval scheme: Ed25519 single-use, expiring tokens with
a first-writer-wins local key store that is safe under concurrent creation, and
a secret-safe audit discipline (\S\ref{sec:sec}).
\item A bounded-adapter execution model that delegates work to external Jekko
and Jeryu engines without surrendering policy, and an ecosystem producer that
backs a live tool-relationship graph (\S\ref{sec:adapters}).
\item A conformance-gated, multi-agent development process enforced by a
regression ratchet at commit, push, and CI time (\S\ref{sec:ratchet}), with an
evaluation of its behavior (\S\ref{sec:eval}).
\end{enumerate}

\section{Background and Related Work}

\textbf{Protocol-governed automation.} Recent agent tooling standardizes how
clients and capability providers exchange requests, but typically leaves
policy, durability, and audit to the host. JMCP keeps that separation but moves
policy enforcement, persistence, and replay into a backend authority that all
clients---user interfaces, chat intake, and adapters---must pass through.

\textbf{Capability-bounded design.} The adapter boundary follows the principle
that the holder of a capability is not the holder of policy. Interfaces and
adapters may request work and display state, but mutation, external calls, and
releases require an approval issued by the backend.

\textbf{Event sourcing and replay.} JPCM borrows from event-sourced systems: a
durable, ordered log of facts from which state is reconstructed. JMCP's
contribution here is operational rather than theoretical---replay is defined as
a reconstruction mechanism that must not repeat side effects, enforced through
idempotency keys and approval references.

\textbf{Static conformance and ratchets.} Quality ratchets that forbid metric
regressions are common for coverage and lint counts. We extend the idea to a
composite static-audit score with categorical ``caps'' and per-rule findings,
and we make it the coordination substrate for multiple autonomous agents
editing a shared tree.

\section{Architecture}
\label{sec:arch}

\subsection{Naming and boundaries}
Three names fix the system boundary. \emph{JMCP} names the system. \emph{JCP/1.0.0}
names the compatibility contract for messages. \emph{JPCM} names the delivery
and replay profile used by the local event bus and by future transports. The
major boundary is between the core runtime and clients of the runtime: user
interfaces and adapters may request work, display state, and submit decisions,
but they do not own policy enforcement. All shared messages become JCP/1.0.0
envelopes before they enter routing, approval, audit, replay, or execution.

\subsection{Workspace topology}
The implementation is a Cargo workspace. Library crates separate the protocol
core (\texttt{jcp-core}: envelope types and Ed25519 signing), domain and store
(\texttt{jmcp-domain}, \texttt{jmcp-store} over embedded SQLite), the
application authority (\texttt{jmcp-app}, \texttt{jmcp-api}), the attention/
needs-attention model (\texttt{jmcp-attention}), conformance
(\texttt{jmcp-conformance}), the typed adapter SDK
(\texttt{jmcp-adapter-sdk}), the three adapters
(\texttt{jmcp-adapter-\{jankurai,jeryu,jekko\}}), and the Telegram approval
channel (\texttt{jmcp-approval-telegram}). Binaries are the daemon
(\texttt{jmcpd}), the control CLI (\texttt{jmcpctl}), and the Rust TUI
(\texttt{jmcp-tui}); the React/TypeScript cockpit is a separate front-end.
Table~\ref{tab:resp} lists V1 responsibilities.

\begin{table}[t]
\centering
\caption{V1 component responsibilities}
\label{tab:resp}
\begin{tabular}{@{}p{0.27\columnwidth}p{0.63\columnwidth}@{}}
\toprule
\textbf{Component} & \textbf{V1 responsibility} \\
\midrule
Rust backend & Protocol validation, approval authority, persistence, replay,
audit, API contracts. \\
SQLite & Durable records for envelopes, approvals, audit entries, adapter
observations, replay checkpoints. \\
JPCM bus & In-process delivery, replay metadata, ordering, idempotency
evidence. \\
Cockpit (React) & Visual inspection, approvals, runtime status, the Jeryu
ecosystem graph. \\
Rust TUI & Terminal operation, recovery, low-dependency inspection. \\
Telegram & Text and voice intake, risk-scored approval, subject to backend
policy. \\
Adapters & Jankurai, Jeryu, Jekko capability under protocol, approval, and
audit constraints. \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Protocol and replay}
A JCP/1.0.0 envelope carries stable identity, version, correlation identity,
source, target or capability, timestamp, payload classification, and message
type. In V1 the in-process event bus is the default JPCM carrier, which makes
local tests deterministic and allows replay without introducing
distributed-system ambiguity before it is needed. Replay rebuilds state from
durable envelopes, approval records, audit facts, and checkpoints;
non-idempotent side effects are skipped, simulated, or recognized as already
committed by idempotency key and approval reference.

\section{Security Model}
\label{sec:sec}

\subsection{Ed25519 approval tokens}
Approval is a first-class control. The backend issues approval tokens signed
with Ed25519 (via \texttt{ed25519-dalek}). A token binds the operation
identity, a single-use nonce, and an expiry; the backend verifies the signature
and rejects tokens that are expired, unknown (forged), or already redeemed. The
test surface proves three properties directly: an expired token is marked
expired, a forged token is unknown, and an approval token is single-use.

The signing key lives in a \texttt{LocalSigner} backed by a local seed file. A
subtle failure mode appears under concurrency: parallel test or worker threads
may each attempt to create the seed file, and a naive create-or-read races
between a half-written file and a second creator, surfacing as
\texttt{AlreadyExists} or ``key file does not contain a 32-byte seed.'' We
resolve this with a \emph{first-writer-wins} construction: each creator writes
its candidate seed to a private per-seed staging file and then attempts an
atomic \texttt{hard\_link} onto the canonical path. Exactly one link succeeds;
losers observe \texttt{AlreadyExists}, discard their staging file, and read the
winner's seed. There are no torn reads because the canonical path only ever
appears fully formed. This made the approval tests deterministic under the
parallel test runner, which had previously passed only when a key file happened
to pre-exist.

\subsection{Secret-safe audit}
Audit records must be useful without becoming secret stores. Secret-like
fields---tokens, passwords, API keys, authorization headers, cookies---are
redacted from logs and audit exports, and raw inbound payload logging is off by
default. Diagnostic records preserve event id, correlation id, actor,
component, operation, decision, timestamp, and diagnostic class. Repository
secrets (for example the Telegram bot token) are kept out of version control
and allow-listed in the secret scanner so the security lane stays both strict
and green.

\subsection{Risk-scored Telegram approval}
Telegram is an approval channel, never a policy replacement. Beyond text
intake, the channel accepts voice notes and attaches a risk score to the
requested operation; higher-risk operations require explicit confirmation
before a token is minted. The backend remains the authority: a Telegram
``approve'' is only honored if it resolves to a valid, unexpired, single-use
token for the correlated operation.

\section{Bounded Adapters and External Execution}
\label{sec:adapters}

Adapters implement a single trait---\texttt{execute(\&WorkOrder)} returning a
vector of \texttt{Evidence}---and are constructed with an injectable client so
that tests run against deterministic stubs while production uses real
transports. The typed \texttt{AdapterError} carries purpose, reason,
common fixes, an optional docs URL, and a repair hint, so failures are
actionable to both operators and agents rather than opaque.

\textbf{Jekko worker engine.} Real work is delegated to Jekko, reached over an
HTTP control endpoint and an OpenAI-compatible fusion endpoint
(\texttt{/v1/chat/completions}), with deeper jobs dispatched through a router
that exposes \texttt{worker\_run} and \texttt{worker\_team} over JSON-RPC. The
adapter maps engine outcomes into JCP evidence and surfaces partial or degraded
results explicitly rather than silently. Crucially, workers are powered by the
external engine, not by the controlling agents; the control plane orchestrates
and audits but does not itself perform the work.

\textbf{Jeryu ecosystem.} The Jeryu adapter exposes an
\texttt{EcosystemSnapshot} of tools and their relationships across repositories,
plus explicit degradation when the producer is unavailable. The snapshot is
serialized in the shape the cockpit consumes directly, which backs a live
operator view of every tool in the ecosystem, how tools relate, the active work
queue, and which systems need attention---rendered in React/TypeScript and kept
current over a streaming channel.

\section{Conformance-Gated Multi-Agent Development}
\label{sec:ratchet}

\subsection{The problem}
JMCP V1 was developed by two autonomous agents editing a shared working tree
concurrently. In that setting, conformance regressions are the default risk:
one agent's edit can re-introduce a finding another removed, or a mid-edit tree
can be committed below an already-accepted quality bar. We make non-regression
a mechanical property.

\subsection{The auditor and the score}
A static auditor (Jankurai) scores the repository on a 0--100 scale, applies
categorical \emph{caps} (for example, ``no security lane on a high-risk repo''
or ``release-readiness gap''), and emits per-rule \emph{findings}
(\texttt{HLT-}$xxx$). The audit is the to-do list: each finding carries its own
problem statement and suggested fix, and caps map to whole categories of work
(CI hardening, dead-language markers, typed errors, UX-QA lanes, supply-chain
evidence, and so on).

\subsection{The ratchet}
We define a compact, accepted \emph{floor}: a summary
$\{\textit{score}, \textit{raw\_score}, \textit{caps\_applied},
\textit{rule\_counts}\}$ stored in the repository. A comparator re-audits the
current tree and \textbf{rejects} the change if any of the following worsen
relative to the floor:
\[
\textit{score}\downarrow \;\lor\;
|\textit{caps}|\uparrow \;\lor\;
\textstyle\sum \textit{findings}\uparrow .
\]
A change that alters \emph{which} cap is present without raising the count or
lowering the score is reported as a note (``fix this next'') but allowed, so
net improvements are never blocked. The comparator is invoked from
\texttt{ops/git-hooks/pre-commit} and \texttt{pre-push} (activated by pointing
\texttt{core.hooksPath} at the repository's hooks) and from a CI lane, so the
gate holds locally and---authoritatively---in continuous integration. The floor
only ratchets \emph{up}, via an explicit \texttt{--accept} after a clean,
improved audit.

\subsection{Reproducibility of the floor}
A practical subtlety is that the audit must be evaluated on the
\emph{committed} state, not a developer's working tree: generated and
git-ignored artifacts (for example a UX-QA report) can suppress caps that a
clean checkout would show. We therefore accept the floor from a fresh,
detached checkout of \texttt{HEAD}, so the bar that every agent's clean clone
sees is the bar the ratchet enforces. This avoids a floor that silently depends
on uncommitted state.

\subsection{Coordination protocol}
The two agents coordinate through an append-only chat log and a claimable work
plan in the repository. Work is scoped (each agent edits disjoint paths),
landed inside explicit commit windows, and---because the ratchet runs on every
commit---neither agent can regress conformance while fixing it. The audit is
re-run after each batch and the floor is bumped as the score improves toward the
target of $\geq 85$ with zero caps and zero high-severity findings.

\section{Reproducibility Standard}

The V1 test bar is CI--local parity. The same embedded SQLite and in-process
JPCM defaults that developers use are exercised by CI; remote services are
excluded from the default proof path; Telegram and adapter integrations provide
local fakes or fixtures. The same lane scripts (\texttt{ops/ci/*.sh}) are
invoked by the local runner and by each GitHub workflow, so ``thin'' workflows
call shared scripts rather than duplicating logic. Required proof lanes are
unit, integration-local, protocol, security, operations, and reproducibility.
Reproducibility tests prove that fixtures, declared generated zones, and
documented commands suffice to recreate the evidence.

\section{Evaluation}
\label{sec:eval}

\subsection{Implementation scale}
The Rust workspace is twelve library crates and three Rust binaries plus a
React/TypeScript cockpit and a Rust TUI: approximately 6{,}800 lines of Rust
under 83 automated tests. Three GitHub workflows (CI, security, and the
Jankurai audit) run the same lane scripts as the local runner, and all three
are green on the committed mainline.

\subsection{Approval correctness under concurrency}
Before the first-writer-wins key store, the approval tests
(\emph{expired-token}, \emph{forged-token}, \emph{single-use}) passed only when
the seed file pre-existed and raced otherwise; after the fix they were green
across repeated parallel stress runs and in CI. This is the concrete payoff of
treating the local profile as the audited profile: a concurrency bug in the
\emph{key store} would have been a security bug in \emph{approval}.

\subsection{Ratchet behavior}
We exercised the ratchet two ways. A deliberately regressing change (re-adding a
dead-language marker, or any edit that raises the cap or finding count) is
rejected at commit time with the exact offending cap/finding printed. A clean,
net-improving change passes, and the floor is then accepted upward. Evaluating
the floor on a detached checkout of \texttt{HEAD} (rather than a dirty tree)
exposed a two-cap discrepancy between the committed state and an uncommitted
batch, confirming the need to seed the floor from committed state. The CI
ratchet lane runs in advisory mode so that the audit job stays green while the
local hooks enforce; promoting the lane to required is a one-line change once
the score clears the target.

\section{Threats to Validity}

The single-machine profile is deliberately narrow; distributed transport,
multi-tenant isolation, and high-availability replay are explicitly out of V1
scope, and the JPCM in-process bus assumes a single ordering authority. The
ratchet enforces \emph{non-regression}, not absolute quality: a repository can
sit at a low-but-stable floor, so the score target must be driven up
separately. Categorical caps compress heterogeneous work into one count, which
can mask a hard finding behind an easy one; we mitigate this by also comparing
total findings and by reporting changed-but-not-worse cap sets. Finally, the
external engines (Jekko, Jeryu) are evaluated through deterministic stubs in CI;
their live behavior is exercised out of band.

\section{Conclusion}

JMCP V1 is a local system with production-shaped discipline: protocol-first
messages, durable state, replayable events, cryptographic single-use approvals,
bounded adapters over real external engines, secret-safe audit, and
deterministic verification at CI--local parity. Its construction contributes a
conformance-gated, multi-agent process in which a regression ratchet---at
commit, push, and CI---makes it mechanically impossible to lower the
repository's audited quality while still permitting every net improvement. The
result is deliberately small enough to run on one machine and strict enough to
serve as a trustworthy foundation for expansion beyond the single-machine
profile.

\begin{thebibliography}{9}
\bibitem{ed25519} D.~J. Bernstein, N.~Duif, T.~Lange, P.~Schwabe, and B.-Y.
Yang, ``High-speed high-security signatures,'' \emph{Journal of Cryptographic
Engineering}, vol.~2, no.~2, pp.~77--89, 2012.
\bibitem{eventsourcing} M.~Fowler, ``Event Sourcing,'' martinfowler.com, 2005.
\bibitem{sqlite} D.~R. Hipp et~al., ``SQLite,'' \url{https://sqlite.org}.
\bibitem{telegrambot} Telegram, ``Bot API,''
\url{https://core.telegram.org/bots/api}.
\bibitem{ci} P.~M. Duvall, S.~Matyas, and A.~Glover, \emph{Continuous
Integration: Improving Software Quality and Reducing Risk}. Addison-Wesley,
2007.
\end{thebibliography}

\end{document}
